---
title: "La Haute Borne Data Exploration"
output:
  html_notebook: default
  github_document: default
---
# Data Exploration Project
```{r setup, include = FALSE}
library(ggplot2)
library(viridis)
```

Data:
```{r}
knitr::kable(
  data[1:5, 1:8], 
  caption = "The data table."
)
```

These are measurments every 10 mins, and so there should be 144 measurments per day.
Check that we are dealing with just year 2017, with `r unique(year(data$Date_time))`.
Count how many measurements were made each month
```{r}
data %>% 
    group_by(month, Turbine) %>%
    count() #tally() 
```



## Missing values

How many missing values?
```{r}
sum(is.na(data$P_avg))
```
missing values per col:
```{r}
na_count <-sapply(data, function(y) sum(is.na(y)))
```
this tells us which vars have a lot of data missing.
```{r}
na_log <- na_count < 10000 # creats a logical vector for subsetting
```
most vars have a few hundred obs missing, however there are 5 vars with 
50% or more missing obs, and one with around 10,000 missing obs.
```{r}
table(na_log) # how many have at least 10000 obs missing
data <- data[, na_log] # remove those vars
sapply(data, function(y) sum(is.na(y))) # recheck
```
these vars have few hundred missing obs.
```{r}
data %>%
    filter(is.na(Ba_avg)) %>% # to show just NA cols
    sapply(function(y) sum(is.na(y))) # this whole block of data is missing
```
```{r}
data %>%
    filter(is.na(Ba_avg)) %>%
    select(c(1:5)) %>%
    summary()
```
This show us it's not just one turbine etc.
Next let's count the NAs per day
```{r}
data %>%
    filter(is.na(Ba_avg)) %>%
    group_by(jday, Turbine) %>%
    summarise(count = n()) %>%
    ggplot(aes(jday, count, colour = Turbine)) +
    geom_line()
```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+L*.

## Special Values
 
Now look at times when the turbine stopped rotating. First look at the distribution of avg rotation speed.
```{r}
summary(data$Rs_avg) # note: there are 843 missing values in this set.
```
```{r}
ggplot(data, aes(Rs_avg, colour = Turbine)) +
    geom_freqpoly(binwidth = 0.2) #+ 
    #coord_cartesian(xlim = c(2, 7))
```
This strange distribution can probably be explained by how a turbine is operated.
Are there any values between 2 and 7?
```{r}
data %>% filter(between(Rs_avg, 2, 7)) %>% group_by(Turbine) %>% 
    summarise(count = n())
```
there are some, but not so many, let's look at their dist:
```{r}
ggplot(filter(data, between(Rs_avg, 2, 7)), aes(Rs_avg, colour = Turbine)) +
    geom_freqpoly(binwidth = 0.2) 
```
quite spread out over range. This is prob the slowing down and speeding up to normal operational speed. There is a concentration at rot-speed = 0. This will prob happen when wind is too low or too high, or during maintenance. To answer this let's look at the dist of wind speed for these values
```{r}
summary(data$Ws_avg)
```
It looks like the rotor speed goes to zero mostly because the wind speed is too low.
```{r}
ggplot(data, aes(Ws_avg, ..density.., fill = Rs_avg < 2)) +
    geom_histogram(position = "dodge")
```
This is an important plot!
```{r}
ggplot(filter(data, Rs_avg < 2 & Turbine == "T1"), aes(jday)) +
    geom_histogram(binwidth = 1)
```
are these zero rotor speed in series in time or random, plot duration of zero rotor speed

This next plot shows how many times per day the rotor stopped
```{r}
ggplot(filter(data, Rs_avg < 0.01)) +
    geom_bar(mapping = aes(x = jday, fill = Turbine), position = "dodge")
```


