---
title: "La Haute Borne Data Exploration"
output:
  html_notebook: default
  github_document: default
---
# Data Exploration Project
```{r setup, include = FALSE}
library(tidyverse)
library(viridis)
library(lubridate)
```

Data:
```{r}
knitr::kable(
  data[1:5, 1:8], 
  caption = "The data table."
)
```

These are measurments every 10 mins, and so there should be 144 measurments per day.
Check that we are dealing with just year 2017, with `r unique(year(data$Date_time))`.
Count how many measurements were made each month
```{r}
data %>% 
    group_by(month, Turbine) %>%
    count() #tally() 
```



## Missing values

How many missing values?
```{r}
sum(is.na(data$P_avg))
```
missing values per col:
```{r}
na_count <-sapply(data, function(y) sum(is.na(y)))
```
this tells us which vars have a lot of data missing.
```{r}
na_log <- na_count < 10000 # creats a logical vector for subsetting
```
most vars have a few hundred obs missing, however there are 5 vars with 
50% or more missing obs, and one with around 10,000 missing obs.
```{r}
table(na_log) # how many have at least 10000 obs missing
data <- data[, na_log] # remove those vars
sapply(data, function(y) sum(is.na(y))) # recheck
```
these vars have few hundred missing obs.
```{r}
data %>%
    filter(is.na(Ba_avg)) %>% # to show just NA cols
    sapply(function(y) sum(is.na(y))) # this whole block of data is missing
```
```{r}
data %>%
    filter(is.na(Ba_avg)) %>%
    select(c(1:5)) %>%
    summary()
```
This show us it's not just one turbine etc.
Next let's count the NAs per day
```{r}
data %>%
    filter(is.na(Ba_avg)) %>%
    group_by(jday, Turbine) %>%
    summarise(count = n()) %>%
    ggplot(aes(jday, count, colour = Turbine)) +
    geom_line()
```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+L*.

## Special Values
 
Now look at times when the turbine stopped rotating. First look at the distribution of avg rotation speed.
```{r}
summary(data$Rs_avg) # note: there are 843 missing values in this set.
```
```{r}
ggplot(data, aes(Rs_avg, colour = Turbine)) +
    geom_freqpoly(binwidth = 0.2) #+ 
    #coord_cartesian(xlim = c(2, 7))
```
This strange distribution can probably be explained by how a turbine is operated.
Are there any values between 2 and 7?
```{r}
data %>% filter(between(Rs_avg, 2, 7)) %>% group_by(Turbine) %>% 
    summarise(count = n())
```
there are some, but not so many, let's look at their dist:
```{r}
ggplot(filter(data, between(Rs_avg, 2, 7)), aes(Rs_avg, colour = Turbine)) +
    geom_freqpoly(binwidth = 0.2) 
```
quite spread out over range. This is prob the slowing down and speeding up to normal operational speed. There is a concentration at rot-speed = 0. This will prob happen when wind is too low or too high, or during maintenance. To answer this let's look at the dist of wind speed for these values
```{r}
summary(data$Ws_avg)
```
It looks like the rotor speed goes to zero mostly because the wind speed is too low.
```{r}
ggplot(data, aes(Ws_avg, ..density.., fill = Rs_avg < 2)) +
    geom_histogram(position = "dodge")
```
This is an important plot!
```{r}
ggplot(filter(data, Rs_avg < 2 & Turbine == "T1"), aes(jday)) +
    geom_histogram(binwidth = 1)
```
are these zero rotor speed in series in time or random, plot duration of zero rotor speed

This next plot shows how many times per day the rotor stopped
```{r}
ggplot(filter(data, Rs_avg < 0.01)) +
    geom_bar(mapping = aes(x = jday, fill = Turbine), position = "dodge")
```

let's look at other variables
```{r}
summary(data$Ds_avg) # note: there are 768 missing values in this set.
ggplot(data, aes(S_avg, colour = Turbine)) +
    geom_freqpoly()#binwidth = 0.2)

```

let's get a daily averaged energy output per turbine i.e. group by day and turbine

What is the mean Active Power per month?

Is there a significant Power diff between turbines?
```{r}
ggplot(data, aes(Turbine, P_avg)) + 
    geom_violin()
ggplot(data, aes(P_avg, colour = Turbine)) +
    geom_freqpoly(binwidth = 80)
ggplot(data, aes(P_avg, colour = Turbine)) +
    geom_density()
```

At first look, it looks like there is no significant difference. Let's make this statment more concrete.

Let's look at the distribution of wind direction 
```{r}
ggplot(data, aes(Wa_avg,  colour = Turbine)) +
    geom_freqpoly(binwidth = 20)
```
the distributions clearly show two peaks i.e. there is a lot more wind in those two specific directions. Are these wind directions related to month of the year? Looking at just one turbine:
```{r}
data %>% 
    filter(Turbine == "T1") %>%
    ggplot(aes(Wa_avg, y = ..density.., colour = factor(month))) +
    geom_freqpoly(binwidth = 20)
```
Inclonclusive. There is no clear pattern between time of year and wind direction. What else could explain these patterns? Outdoor temp?
```{r}
data %>% 
    filter(Turbine == "T1") %>%
    ggplot(aes(Wa_avg, Ot_avg)) + 
    geom_hex() +
    scale_fill_viridis()
```

Perhaps we can see a slight difference here. Very hard to see. Also not conclusive

Then summarise. 
Note: we are taking the avg of avgs. This is the same as the total avg if 
the sample sizes are all the same, in this case we can assume that they are. Otherwise would have to use: `weighted.mean`
```{r}
month_avgP <- data %>% 
    group_by(month, Turbine) %>%  # group by month, and turbine
    summarize(monthlyPavg = mean(P_avg, na.rm = TRUE)) 
# calculate the monthly mean of power
```

avg by day
```{r}
day_avgP <- data %>% 
    group_by(jday, Turbine) %>%  # group by day, and turbine
    summarize(dailyPavg = mean(P_avg, na.rm = TRUE)) 
# calculate the daiy mean of power
```

First we plot monthly avg Power over 2017
```{r}
ggplot(month_avgP, aes(month, month_avg_P)) +
    geom_line(mapping = aes(color = Turbine)) 
```

notice how there is an order in power. Why is there an order in turbine power, constantly throughout the year?

Next we plot daily avg Power over 2017
```{r}
ggplot(day_avgP, aes(jday, dailyPavg)) +
    geom_line(mapping = aes(color = Wind_turbine_name)) 
```

find times when the turbine stopped

what var is most highly correlated with Power?

next look at avg wind speed vs wind direction
```{r}
ggplot(data, aes(Wa_avg, Ws_avg)) +
    geom_point(alpha = 0.3) + #mapping = aes(color = Wind_turbine_name))
    facet_wrap(~ Turbine) 
```

What is special about these points? It's the speeding-up or slowing-down values
```{r}
ggplot(data, aes(Ws_avg, P_avg, colour = Rs_avg < 8)) + 
    geom_point(alpha = 0.4) 
```
look at the weird data:
```{r}
data %>% filter(Ws_avg > 12.5 & between(P_avg, 1000, 1500))
normaldata <- summary(filter(data, Ws_avg < 11 & between(P_avg, 1000, 1500)))
normaldata[c(4),] # picks out mean
```

```{r}
ggplot(data, aes(I(Ws_avg^3), P_avg)) + 
    geom_point(alpha = 0.1) + 
    geom_smooth(colour = "red") + 
    theme_bw()
```





